---
title: "Wrangling"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dslabs)
library(tidyverse)
library(readr)
```

## Section 1: Data Import

Testing working directory and file locations.

```{r}

system.file("extdata", package="dslabs")
path <- system.file("extdata", package="dslabs")
list.files(path)

filename <- "murders.csv"
fullpath <- file.path(path,filename)
file.copy(fullpath,getwd())
file.exists(filename)

```
Let's take a look at the first few lines to confirm what kind of file we're dealing with.


```{r}
read_lines("murders.csv", n_max=3)
dat <- read_csv(filename)
view(dat)

```
Importing data files from the internet.

```{r}
url <- "https://raw.githubusercontent.com/rafalab/dslabs/master/inst/extdata/murders.csv"
dat <- read_csv(url)
download.file(url, "murders.csv")

getwd()
filename <- "murders.csv"
path <- system.file("extdata", package = "dslabs")

file.copy(file.path(path, "murders.csv"), getwd())

setwd("data")
file.copy(file.path(path, filename), getwd())

file.copy(file.path(path, "murders.csv"), file.path(getwd(), "data"))

file.location <- file.path(system.file("extdata", package = "dslabs"), "murders.csv")
file.destination <- file.path(getwd(), "data")
file.copy(file.location, file.destination) 
```

Assignment

```{r}
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/wdbc.data"

tmp_filename <- tempfile()
download.file(url, tmp_filename)
read_lines(tmp_filename, n_max=3)
dat <- read_csv(tmp_filename,col_names=FALSE)
file.remove(tmp_filename)


```


## Section 2: Tidy Data

In the Tidy Data section, you will learn how to convert data from a raw to a tidy format.

This section is divided into three parts: Reshaping Data, Combining Tables, and Web Scraping. There are comprehension checks at the end of each part.

```{r}
data(gapminder)

# create and inspect a tidy data frame
tidy_data <- gapminder %>% 
  filter(country %in% c("South Korea", "Germany")) %>%
  select(country, year, fertility)
head(tidy_data)

# plotting tidy data is simple
tidy_data %>% 
  ggplot(aes(year, fertility, color = country)) +
  geom_point()

# import and inspect example of original Gapminder data in wide format
path <- system.file("extdata", package="dslabs")
filename <- file.path(path,  "fertility-two-countries-example.csv")
wide_data <- read_csv(filename)
select(wide_data, country, `1960`:`1967`)

```











